---
title: "Exploring SafaApp data"
author: "Eduard Bukin"
output:
  html_document: default
  html_notebook: default
---

Below, I attempt to explor data on the usage of the SAFA App along with the rates of app's instalation and deletion by countries. 

# Setup and Loading data

Previously, data on the app usage was downloaded in the bulk from the the google app portal and was farvested with the R code.

```{r error=FALSE,warning=FALSE,message=FALSE}
# Packages
library(tidyverse)
library(stringr)
library(lubridate)

# Loading app related data
load("allHarvestedAppData.Rdata")
```

# Data exploration

Cleaning data:

```{r}
vars <- 
  appData %>% 
  filter(fileName == "country") %>% 
  map_lgl(.x = ., .f = function(x) {!all(is.na(x))})

countryData <-
  appData[vars] %>%
  mutate(Date = ymd(Date)) %>% 
  filter(fileName == "country") %>%
  select(Date, Country, Daily.Device.Installs, Daily.Device.Uninstalls) %>% 
  distinct() 
```

Looking at the total number of the apps instalation and deletions over the year:  
`Daily.Device.Installs` means how many time the app was installed on one device  
`Daily.Device.Uninstalls` means how many times the app was deleted from one device  
Both these numbers may include repetative installation, when the app was reinstalled on the same device.

```{r}
countryData %>% 
  group_by(Year = year(Date)) %>% 
  summarise_at(c(3,4), sum)
  
```

Make plot by countries.

```{r}
countryData %>% 
  group_by(Country, Year = year(Date)) %>% 
  summarise_at(c(3,4), sum) %>% 
  group_by(Country) %>% 
  mutate(SummOf= sum(Daily.Device.Installs)) %>% 
  ungroup() %>% 
  arrange(desc(SummOf), Country, Year) %>% 
  knitr::kable()

countryData
  
```










