---
title: "Exploring SafaApp data"
author: "Eduard Bukin"
output:
  pdf_document: default
---

Below, I attempt to explore data on the usage of the SAFA App along with the rates of app's installation and deletion by countries. 

# Setup and Loading data

Previously, data on the app usage was downloaded in the bulk from the the Google app portal and was harvested with the R code.

```{r error=FALSE,warning=FALSE,message=FALSE}
# Packages
library(tidyverse)
library(stringr)
library(lubridate)
library(forcats)
devtools::install_github('UNFAOstatistics/gisfao')
library(gisfao)
library(tmap)
library(countrycode)

# Loading app related data
load("data/allHarvestedAppData.Rdata")
```

# Data exploration

```{r}
vars <- 
  appData %>% 
  filter(fileName == "country") %>% 
  map_lgl(.x = ., .f = function(x) {!all(is.na(x))})

countryData <-
  appData[vars] %>%
  mutate(Date = ymd(Date)) %>% 
  filter(fileName == "country") %>%
  select(Date, Country, Daily.Device.Installs, Daily.Device.Uninstalls) %>% 
  distinct() %>% 
  gather(Var, Val, Daily.Device.Installs, Daily.Device.Uninstalls) %>% 
  mutate(Year = as.factor(as.character(year(Date))),
         Country = as.factor(Country),
         Var = str_extract(Var, "[^.]*$"))  
```

Looking at the total number of the apps installation and deletions over the year:  
`Installs` means how many time the app was installed on one device  
`Uninstalls` means how many times the app was deleted from one device  
Both these numbers may include repetitive installation, when the app was re installed on the same device.

```{r}
countryData %>% 
  group_by(Year, Var) %>% 
  summarise_at("Val", sum) %>% 
  ungroup() %>% 
  mutate(Year = fct_reorder(Year, as.integer(Year), .desc = TRUE)) %>% 
  ggplot(aes(x = Var, y = Val, fill = Year)) + 
  geom_col(aes(fill = Year)) +
  geom_text(aes(label = Val), position = position_stack(vjust = 0.5)) +
  xlab("") +
  ylab("Number of installs / uninstalls")
  
```

# Number of installs and uninstalls by countries and years 

Please note that for this plot we filtered only those countries, which had more than 10 installs of the SAFAApp. 

```{r}
countryData %>% 
  group_by(Country, Year, Var) %>% 
  summarise_at("Val", sum) %>% 
  group_by(Country) %>% 
  mutate(SummOf = sum(Val)) %>% 
  ungroup() %>% 
  arrange(desc(SummOf), Country, desc(Year), Var) %>% 
  filter(SummOf > 10) %>% 
  mutate(Countries = fct_reorder(as.factor(str_c(Country, " ", Var)), SummOf, .desc = TRUE),
         Year = fct_reorder(Year, as.integer(Year),.desc = TRUE )) %>% 
  ggplot(aes(x = Countries, y = Val, fill = Year))  + 
  geom_col() +
  geom_text(aes(label = Val), position = position_stack(vjust = 0.5)) +
  # geom_bar(stat = "identity", position = "stack") +
  coord_flip() + 
  theme(axis.text.y=element_text(hjust=0))+
  xlab("") +
  ylab("Number of installs / uninstalls")
  
```

```{r echo=FALSE}
installsNumber <-
  countryData %>% 
  filter(Var == "Installs", Val != 0) %>% 
  select(Country) %>%  
  distinct() %>%  
  nrow()

currentInstallsNumber <-
  countryData %>% 
  group_by(Country, Var) %>% 
  summarise(Val = sum(Val)) %>% 
  ungroup() %>% 
  spread(Var, Val) %>% 
  mutate(dif = Installs - Uninstalls) %>% 
  filter(dif > 0 ) %>% 
  select(Country) %>% 
  distinct() %>% 
  nrow()

```


Number of countries, where the App was installed at some point is `r installsNumber`.  

Number of countries, where the app is currently installed is `r currentInstallsNumber`.

Top 10 countries, where the app os used:

```{r}
countryData %>% 
  group_by(Country, Var) %>% 
  summarise_at("Val", sum) %>% 
  ungroup() %>% 
  filter(Var == "Installs", Val > 0) %>% 
  arrange(desc(Val)) %>% 
  top_n(10, Val) %>% 
  mutate(Country = str_trunc(countrycode(Country, "iso2c", "country.name.en"), width = 30)) %>% 
  ggplot(aes(x = fct_reorder(Country, Val), y = Val, fill = ""))  + 
  scale_fill_manual(values = c("skyblue"), name = "")+
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Val), position = position_stack(vjust = 0.5)) +
  coord_flip()+ 
  theme(axis.text.y=element_text(hjust=0))+
  xlab("") +
  ylab("Number of installs by country out of top 10")
  
```


# Mapping SAFA usage data against the world map

The numbers below indicate nymber of installes of SAFA App on the various android smartphones.

```{r error=FALSE,warning=FALSE,message=FALSE}
# Loading FAO world shape data file
data(fao_world)

# Grid lines
data(graticule)

# Actual data with countrycodes
ActualData <- 
  countryData %>% 
  group_by(Country, Var) %>% 
  dplyr::summarise(Val = sum(Val)) %>% 
  ungroup() %>% 
  spread(Var, Val) %>% 
  dplyr::filter(Country != "", !is.na(Country)) %>% 
  dplyr::mutate(FAOST_CODE = countrycode(Country, "iso2c", "fao")) %>% 
  # dplyr::filter(is.na(FAOST_CODE)) %>% 
  dplyr::mutate(FAOST_CODE = ifelse(as.character(Country) == "PS", 299, FAOST_CODE),
         FAOST_CODE = ifelse(as.character(Country) == "PR", 177, FAOST_CODE),
         FAOST_CODE = ifelse(as.character(Country) == "RE", 182, FAOST_CODE)) %>% 
  right_join(tibble(FAOST_CODE = fao_world$FAO_CODE), by = "FAOST_CODE") %>% 
  select(-Country)

# Adding data variables
fao_world$Installs <- ActualData$Installs
fao_world$Uninstalls <- ActualData$Uninstalls

# Plotting the map
SafaMap <- 
  tm_shape(fao_world, projection="robin") +
  # tm_borders() +
  tm_polygons("Installs", palette="Blues", 
              style = "kmeans", 
              contrast = 1, 
              id="name", 
              title="SAFA App usage",
              legend.z = 1000) +
  tm_grid(projection="longlat", labels.size = 0, alpha = .8) + 
  tm_style_white() + 
  tm_format_World()

# Savingthe map in the file
save_tmap(SafaMap, "output/Safa_World_map.png", width=1920, height=1080)

# Plotting the map
SafaMap

```










